---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [The People Streaming Church Aren’t Who You Think](https://thescoop.org/religion/reverse-engineering//streaming-church.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(refinr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

This data is collected from the 2023-2024 Religious Landscape Study, which Pew Research conducted. My basic dictionary can be found below. The main data points I will be using are highest degree (EDUCREC), attendance in person (ATTNDPERRLS), attendance online (ATTNDONRLS), decade of birth (BIRTHDECADE), and totaly family income (INC_SDT1).

There are three main caveats I see with this data. For one, the average weight for is around 7,000, which means that one person's answer accounts for 7,000 people. However, it's clear some outliers exist, as the median weight is around 4,000. Still, this is something to keep in mind, as the polling seems to be heavily skewed towards one demographic. There is also no distinction as to what counts as "very old" in the dataset, as that is something the author decided for himself. With no guideline, I will consider anyone born before 1960 as "very old," and will adjust accordingly if I my findings are off. Finally, there is not exact birth years and only decades. This makes aggregating data based on age incredibly difficult, especially when one of my questions

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

RLS <- read_csv("data/RLS.csv.zip")

# Clean required data and prepare for analysis if needed. 

###RLS <- RLS |>
  ###mutate(
   ### highest_degree = case_when(
     ### EDUCREC == 1 ~ "High school or less",
     ### EDUCREC == 2 ~ "Some college/Associate's degree",
     ###EDUCREC == 3 ~ "College Graduate",
     ###EDUCREC == 4 ~ "Master's degree or higher",
     ### EDUCREC == 99 ~ "No Answer"
   ### )
 ### )

###RLS <- RLS |>
  ###mutate(
   ### attendance_in_person = case_when(
    ###  ATTNDPERRLS == 1 ~ "More than once a week",
    ###  ATTNDPERRLS == 2 ~ "Once a week",
    ###  ATTNDPERRLS == 3 ~ "Once or twice a month",
    ###  ATTNDPERRLS == 4 ~ "A few times a year",
     ### ATTNDPERRLS == 5 ~ "Seldom",
     ### ATTNDPERRLS == 6 ~ "Never",
     ### ATTNDPERRLS == 99 ~ "Don't know/Refused"
    ###)
 ### )

###RLS <- RLS |> 
  ###mutate (
    ###attendance_online = case_when(
    ###  ATTNDONRLS == 1 ~ "More than once a week",
    ###  ATTNDONRLS == 2 ~ "Once a week",
    ###  ATTNDONRLS == 3 ~ "Once or twice a month",
     ### ATTNDONRLS == 4 ~ "A few times a year",
     ### ATTNDONRLS == 5 ~ "Seldom",
     ### ATTNDONRLS == 6 ~ "Never",
     ### ATTNDONRLS == 99 ~ "Don't know/Refused"
    ###)
  ###)

###RLS <- RLS |>
  ###mutate(
    ###birth_decade = case_when(
      ###BIRTHDECADE == 1 ~ "1940s or earlier",
      ###BIRTHDECADE == 2 ~ "1950s",
     ###BIRTHDECADE == 3 ~ "1960s",
      ###BIRTHDECADE == 4 ~ "1970s",
      ###BIRTHDECADE == 5 ~ "1980s",
      ###BIRTHDECADE == 6 ~ "1990s",
      ###BIRTHDECADE == 7 ~ "2000s",
      ###BIRTHDECADE == 99 ~ "No answer"
    ###)
  ###)

###RLS <- RLS |>
  ###mutate(
    ###total_family_income = case_when(
     ### INC_SDT1 == 1 ~ "Less than $30,000",
     ### INC_SDT1 == 2 ~ "$30,000 to less than $40,000",
     ### INC_SDT1 == 3 ~ "$40,000 to less than $50,000",
     ### INC_SDT1 == 4 ~ "$50,000 to less than $75,000",
      ###INC_SDT1 == 5 ~ "$75,000 to less than $100,000",
     ### INC_SDT1 == 6 ~ "$100,000 to less than $125,000",
     ### INC_SDT1 == 7 ~ "$125,000 to less than $150,000",
     ### INC_SDT1 == 8 ~ "$150,000 or more",
     ### INC_SDT1 == 99 ~ "Don't know/Refused"
    ###)
 ### )

RLS_curated <- RLS |>
  select(P_SUID, WEIGHT, EDUCREC, ATTNDPERRLS, ATTNDONRLS, BIRTHDECADE, INC_SDT1)

RLS_curated |>
  summarise(
    min_weight = min(WEIGHT),
    median_weight = median(WEIGHT),
    mean_weight = mean(WEIGHT),
    max_weight = max(WEIGHT)
  )
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: ["Among very old Americans, the ones who are least likely to go to church are those who stopped at high school."]
* **Analysis summary**: [I was able to confirm this finding. Based on weighted percentages, around 35% of American adults who were born in the 1950s or earlier completed only a high school degree or less, more than any other answer. I also ran a test to see what the top attendance rates were within this specific demographic, and some of the top results are "seldom," and "never," and "a few times a year." These answers make up over 60% of the responses.]

```{r}
# Put code to reverse engineer sentence here
RLS_curated_age <- RLS_curated |>
  mutate(is_old = BIRTHDECADE < 3)

RLS_curated_age_educ <- RLS_curated_age |>
  filter(is_old == TRUE) |>
  filter(EDUCREC == 1)

RLS_curated_age_educ |>
  group_by(attendance_in_person) |>
  summarize (
    count = n()
  ) |>
  arrange(desc(count))

RLS_curated_age |>
  summarise(
    hs_weight = sum(WEIGHT[EDUCREC == 1]),
    total_weight = sum(WEIGHT)
  ) |>
  mutate(hs_pct = hs_weight/total_weight * 100)

RLS_curated_age |>
  summarise(
    somecol_weight = sum(WEIGHT[EDUCREC == 2]),
    total_weight = sum(WEIGHT)
  ) |>
  mutate(somecol_pct = somecol_weight/total_weight * 100)

RLS_curated_age |>
  summarise(
    col_weight = sum(WEIGHT[EDUCREC == 3]),
    total_weight = sum(WEIGHT)
  ) |>
  mutate(col_pct = col_weight/total_weight * 100)

RLS_curated_age |>
  summarise(
    masters_weight = sum(WEIGHT[EDUCREC == 4]),
    total_weight = sum(WEIGHT)
  ) |>
  mutate(masters_pct = masters_weight/total_weight * 100)

RLS_curated_age |>
  summarise(
    misc_weight = sum(WEIGHT[EDUCREC == 2]),
    total_weight = sum(WEIGHT)
  ) |>
  mutate(misc_pct = misc_weight/total_weight * 100)



  

  
# Display results of code below this codeblock

Seldom	650			
Once a week	648			
Never	470			
A few times a year	426			
More than once a week	281			
Once or twice a month	192			
Dont know/Refused	37	

```

### Sentence 2

* **Sentence text**: ["I will write this until I’m blue in the face: education is positively related to religious attendance in the United States."]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here



# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Who is the most likely to show up on Sunday? It’s college-educated folks who make $50K–$100K per year.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

RLS_curated_inc <- RLS_curated |>
  mutate (is_in_range = INC_SDT1 >= 4 & INC_SDT1 <= 5) |>
  summarise(
    range_weight = sum(WEIGHT[is_in_range == TRUE]),
    total_weight = sum(WEIGHT)
  ) |>
  mutate(range_pct = range_weight / total_weight * 100)

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-