---
title: "lab_03"
author: "Derek Willis"
date: "2026-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TASK: Load libraries you'll need to load and clean the data

```{r}
# Turn off scientific notation
options(scipen=999)


```

## TASK: Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using `janitor`, change the columns so that they are lower-cased and any spaces are replaced with underscores. You also will want to make it so that you can find the most records possible when using filter(), which means making the records look as consistent as possible in terms of their formatting.

```{r}
# Load in the Maryland state agency payments data

# Load in the Maryland grants and loan data
```

## Answer questions

Q1: You want to know how much money state agencies have paid to religious organizations in a particular denomination, but to do that you'll need to find the variations of how that denomination's organizations might be listed. Pick a denomination or faith and write code to find rows where the vendor name contains the name you're searching for, starting with a simple word or phrase and then getting more complex if needed. NOTE: you need to pick one that has at least some payments!

HINT: you should use filter() just once, and inside your filter you want to look for the variations one at a time. Think "this version OR that version."

When you have found as many as possible, save the result to a new dataframe called \`rel_orgs\`\`. Write the denomination or faith, and how many records you found.

A1:

```{r}
```

Q2: Using that `rel_orgs` dataframe, write code to calculate the total amount of payments to those religious organizations for each fiscal year. Which fiscal year had the largest total amount? Are there any aspects of the results that stand out to you?

A2:

```{r}
```

Q3. You're interested in examining state payments (not grants or loans) to Catholic Charities organizations. Use `filter()` and `str_detect()` to find those payments and save them to a new dataframe. Then look carefully at that list - what cleaning will it require to ensure that you can standardize the vendor names? Next, write a plan - not code for this step - describing how you would clean up the records so that the end result is that you'd have each unique vendor listed only once. You *can* write code for this, but what's required is a step-by-step process.

A3.

```{r}
```

Q4: You want to the leading recipients of state grant money given for the purpose of preventing hate crimes. Filter the data to find grants that fit that description; you'll want to make sure that you locate all of them. Then calculate the total amount of money for each grantee. Write a sentence or two describing the top 20 grantees and identify the most interesting one(s) to you. You can do some research about them to back up your answer.

A4:

```{r}

```

Q5: Using the Maryland grants and loan data, write code to find grant recipients of the African American Heritage Preservation Program (AAHPP) and produce a list of grantees along with the total amount they received. Describe which of the top 20 recipients are churches or church-related (you may need to do some research to figure that out). Which of those grantees are most interesting to you?

A5:

```{r}
```

Q6: Using the denomination that you are collecting church information on, write code to find any payments to those organizations that are listed in the grants and loan data. How many records do you find? What's the most interesting payment you find in the data and why?

A6: 

```{r}
```

