---
title: "lab_07"
author: "Derek Willis"
date: "2025-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(tidycensus)

```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key only if you haven't already. You won't need to install or overwrite it.

```{r echo=FALSE}
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2021 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2021 5-year ACS (which is called "acs5") and save them to a dataframe

```{r}
acs5 <- load_variables(2021, "acs5", cache = TRUE)
```

## Answer questions

**Q1** What is the Maryland county with the highest percentage of the population with a bachelor's degree or higher, according to the ACS 2021 5-year estimates? The variable you want is `DP02_0068P` from the ACS Data Profile tables — it's the pre-calculated percentage of the population 25 and over with a bachelor's degree or higher.

**Howard County has the highest percentage with 63.6%**

```{r}
bd21 <- get_acs(geography = "county", variables = "DP02_0068P", state = "MD", year = 2021)|>
arrange(desc(estimate))
```

**Q2** Now let's join that data with U.S. Religion Census data. Load the `usreligioncensus_county_2020.csv` file and filter it for Maryland counties. Then join it to your education data. Look at the top 5 counties by educational attainment and the bottom 5. How do their religious adherence rates compare? Write a sentence describing what you see.

**A2**The top 5 counties by educational attainment are Howard County, Montgomery County, Anne Arundel County, Frederick County, and Talbot county. The top 5 lowest are Somerset County, Caroline County, Dorchester County, Allegany County and Washington County. Generally, the percent of all adherents tends to be lower the higher the county's educational attainment is. 

```{r}
census_2020 <- read_csv("usreligioncensus_county_2020.csv") 

md_census_2020 <- census_2020 |>
  filter(`State Name` == "Maryland")

education_and_census <-md_census_2020 |>
   inner_join(bd21, join_by(`FIPS`== `GEOID`)) |>
  arrange(desc(estimate))

```

**Q3** Using your joined Maryland county data, find the 5 most and 5 least religiously diverse counties based on the Diversity Index Value. Show the largest religious body by number of adherents and the median household income (you'll need to get income from the ACS and join it in) for those counties. What patterns do you notice about which religious groups dominate in more vs. less diverse counties?

**A3** The five most religiously diverse counties are Worcester County, Baltimore city, Prince George's County, Montgomery County, and Talbot County. The five least are Somerset County, Cecil County, Anne Arundel County, Harford County and Dorchester County. The largest religious body by number of adherents is the Catholic Church in Montgomery County. Its median household income is $117345. Even in religiously diverse counties, the Catholic Church is primarily the most dominant.

```{r}
median_income <- get_acs (geography = "county", variables = "B19013_001", state = "MD", year = 2021)|>
  arrange(desc(estimate))

md_census_2020 <- md_census_2020 |>
  inner_join(median_income, join_by(`FIPS` == `GEOID`))

```

**Q4** Is there a relationship between wealth and the number of religious congregations in a community? Using your joined data that includes median household income, look at the "Congregations Per 100,000 Population" column. Arrange the counties by congregations per capita. Do the counties with the most congregations per capita tend to be wealthier or less wealthy? What about the counties with the fewest? Write a sentence describing the pattern.

**A4** Generally, it looks like there are more congregations per 100,000 population in counties with lower median household incomes. In contrast, there are less congregations per 100,000 population in counties with larger median household incomes. The pattern I am seeing is that less wealthy counties have more congregations. 

```{r}
md_census_2020 |>
  arrange(desc(`Congregations Per 100,000 Population`))
```

**Q5** Let's compare religion and income across all states. Get state-level median household income from the ACS and join it with the state-level religion census data (`usreligioncensus_state_2020.csv`). Remember you'll need to make the FIPS codes match in format, as we did in the pre_lab.

Arrange the states by adherence rate. Among the 10 states with the highest adherence rates, what is the average median household income? Among the 10 states with the lowest adherence rates, what is the average? What regional patterns do you notice in each group?

**A5** Among the 10 highest states, the average median income is $59625.50. Among the 10 lowest states, the average median income is $51387.60. For the highest adherence rate, it is mostly deep south/southern/midwest, with the outlier being Utah, but that can easily be explained by the high Mormon population. On the lower end, it is mostly the east and west coast. 

```{r}
religion_state <- read_csv("usreligioncensus_state_2020.csv") |>
  clean_names()

religion_county <- read_csv("usreligioncensus_county_2020.csv") |>
  clean_names()

median_income <- get_acs(
  geography = "state",
  variables = "B19013_001",
  year = 2020,
  survey = "acs5"
)

median_income <- median_income |>
  select(GEOID, NAME, estimate) |>
  rename(median_income = estimate)

religion_state_clean <- religion_state |>
  mutate(
    state_code = str_pad(
      as.character(state_code),
      width = 2,
      pad = "0"
    )
  )

religion_state_income <- median_income |>
  inner_join(
    religion_state_clean,
    join_by(GEOID == state_code)
  )

```

**Q6** Your own question: Come up with your own question that combines Census API data with U.S. Religion Census data. Your question should:

-   Use at least one ACS variable we haven't used in previous questions
-   Involve a join between Census and religion data
-   Include grouping, ranking, or comparison across geographic areas

Write your question, show your code, and provide an answer with interpretation.

**A6**

Your question: How high are the adherence rates for the congregations with the highest population of citizens with hispanic or latino origin?

```{r}
hispanic_latino <- get_acs (geography = "county", variables = "B03003_001", state = "MD", year = 2021)|>
  arrange(desc(estimate))

md_census_2020 <- md_census_2020 |>
  inner_join(hispanic_latino, join_by(`FIPS` == `GEOID`))
```

Your answer and interpretation: The answer is that the top 5 counties are Montgomery County, Prince George's County, Baltimore County, Baltimore City, and Anne Arundel County, which have adherence rates of 53.42%, 50.04%, 34.66%, 51.33%, and 33.86%. Generally, it's safe to gather from this that congregations with high rates of hispanic/latino citizens tend to have a higher percent of adherents. Interesting to note for all of these is that, with the exception of Baltimore city, all counties have the Catholic church as the largest body by number of adherents. In fact, 9 out of the top 10 all are the Catholic church. 
